---
title: "P8105_HW1_rs4788"
output: html_document
date: "2025-09-19"
---


## Problem 1

```{r}
library("moderndive")
data("early_january_weather")
```


#Description of the dataset:

The dataset contains meteorological data for the first half of January 2013. 
Key variables in this dataset include: Time indicators: `r paste(names(early_january_weather)[2:5], collapse = ",")`,time_hour; Meteorological measurements: `r paste(names(early_january_weather)[6:14], collapse = ",")`.
The mean temperature recorded in this period was `r round(mean(early_january_weather$temp))`°F, 
with humidity averaging `r round(mean(early_january_weather$humid))`%.

The size of the dataset is `r nrow(early_january_weather)` rows 
  and `r ncol(early_january_weather)` columns.
  
The mean temperature is `r mean(early_january_weather$temp)`℉.



#Make a scatterplot:

```{r}
library(ggplot2)
scatter_plot <- ggplot(early_january_weather, 
                       aes(x = time_hour, y = temp, color = humid))+ 
            geom_point()+
              labs(title = "Temperature Variation over Time by Humidity Level",
               x = "Time Hour", y = "Temperature (F)",color = "Humidity") 
print(scatter_plot)


ggsave("temperature_vs_time_scatterplot.png", plot = scatter_plot, 
       width = 10, height = 6)

```
The scatter plot indicates that temperatures showed an overall upward trend in January. Daily fluctuations within the scatter plot reflect temperature variations caused by day-night cycles. Humidity changes are somewhat correlated with temperature, with higher temperatures generally leading to lower humidity levels.